---
params:
  source_url: "https://walker-data.com/posts/overture-buildings"
title: "walker.data.com"
subtitle: "`r params$source_url`"
execute: 
  echo: true
  eval: false
---
# Setup
```{r library}
#| warning: false
#| message: false
if (!requireNamespace('arrow'))     
  install.packages('arrow', 
    repos = 'https://apache.r-universe.dev')
library(arrow) 
# first load arrow then, _if_ needed, run install_arrow()
arrow_info() |> names()
arrow_info()$capabilities |> sort()
arrow_info()$build_info$cpp_version
arrow_info()$build_info$cpp_compiler
stopifnot(
    arrow_available()
  , arrow_with_s3()
  , arrow_with_parquet()
  , arrow_with_json()
  , arrow_with_dataset()
  , arrow_with_gcs()
  # arrow_with_acero() , arrow_with_substrait()
)
if (!requireNamespace('rdeck')) 
  pak::pak("qfes/rdeck")
# remotes::install_github("qfes/rdeck")
if (!requireNamespace('pacman')) 
  install.packages("pacman")
pacman::p_load(
  sf, dplyr, tigris, rdeck,
  scales, tictoc, lobstr)
options(tigris_use_cache = TRUE)
```
```{r open_dataset}
s3_buildings <- 
  's3://overturemaps-us-west-2/release/2024-05-16-beta.0/theme=buildings?region=us-west-2'
tic('open_dataset')
  # database-like connection to the buildings dataset
  buildings <- open_dataset(s3_buildings)
toc()
lobstr::obj_size(buildings)
```
```{r nrow(buildings)}
tic('nrow(buildings)')
  num_buildings <- nrow(buildings) 
toc()
num_buildings |> prettyNum(big.mark = ',')
num_buildings |> 
  scales::label_number(accuracy=0.1, 
    scale_cut=scales::cut_short_scale())(x = _)
```
## London
+ Wide {{bbox|-0.489|51.28|0.236|51.686}}
+ https://www.openstreetmap.org/export#map=15/51.5150/-0.0912
  + ldn_bb <- c(-0.1061, 51.5227, -0.0763, 51.5072)
+ http://bboxfinder.com/#0.000000,0.000000,0.000000,0.000000
  + -0.120450,51.509507,-0.076247,51.526465
```{r ldn_bbox}
ldn_bbox <- c(-0.1061, 51.5227, -0.0763, 51.5072)
get_buildings <- function(bbox, buildings){
  buildings |>
  # Filter expression not supported for Arrow Datasets
  # collect() |> 
  filter(bbox$xmin > bbox[1],
         bbox$ymin > bbox[2],
         bbox$xmax < bbox[3],
         bbox$ymax < bbox[4]) |>
  select(id, geometry, height) |> 
  collect() |>
  st_as_sf(crs = 4326) |> 
  mutate(height = ifelse(is.na(height), 8, height))
}
ldn_buildings <- 
  ldn_bbox |> 
  get_buildings(buildings)
```
```{r rdeck Ldn}
rdeck(map_style = mapbox_light(), 
      initial_view_state = view_state(
        center = c(-122.4657, 37.7548),
        zoom = 11.3,
        bearing = -60,
        pitch = 76
      )) |> 
  add_polygon_layer(
    data = ldn_buildings, 
    name = "London (central)",
    get_polygon = geometry, 
    get_elevation = height, 
    get_fill_color = scale_color_linear(
      col = height,
      palette = viridisLite::inferno(100, direction = -1)
    ),
    extruded = TRUE, 
    opacity = 0.5)
```


<!-- ## SF -->
<!-- ```{r counties} -->
<!-- #| warning: true -->
<!-- #| message: false -->
<!-- sf_bbox <- counties(state = "CA", cb = TRUE, resolution = "20m") |>  -->
<!--   filter(NAME == "San Francisco") |>  -->
<!--   st_bbox() |>  -->
<!--   as.vector() -->
<!-- sf_buildings <-  -->
<!--   sf_bbox |>  -->
<!--   get_buildings(buildings) -->
<!-- sf_buildings_old <- buildings |> -->
<!--   filter(bbox$xmin > sf_bbox[1], -->
<!--          bbox$ymin > sf_bbox[2], -->
<!--          bbox$xmax < sf_bbox[3], -->
<!--          bbox$ymax < sf_bbox[4]) |> -->
<!--   select(id, geometry, height) |>  -->
<!--   collect() |> -->
<!--   st_as_sf(crs = 4326) |>  -->
<!--   mutate(height = ifelse(is.na(height), 8, height)) -->
<!-- sf_buildings |> identical(sf_buildings_old) -->
<!-- ``` -->
<!-- ```{r rdeck SF} -->
<!-- rdeck(map_style = mapbox_light(),  -->
<!--       initial_view_state = view_state( -->
<!--         center = c(-122.4657, 37.7548), -->
<!--         zoom = 11.3, -->
<!--         bearing = -60, -->
<!--         pitch = 76 -->
<!--       )) |>  -->
<!--   add_polygon_layer( -->
<!--     data = sf_buildings,  -->
<!--     name = "San Francisco", -->
<!--     get_polygon = geometry,  -->
<!--     get_elevation = height,  -->
<!--     get_fill_color = scale_color_linear( -->
<!--       col = height, -->
<!--       palette = viridisLite::inferno(100, direction = -1) -->
<!--     ), -->
<!--     extruded = TRUE,  -->
<!--     opacity = 0.5) -->
<!-- ``` -->
<!-- <!-- ## sydney --> -->
<!-- <!-- ```{r sydney_bbox} --> -->
<!-- <!-- sydney_bbox <- c(150.787520, -34.088237, 151.422667, -33.643521) --> -->

<!-- <!-- sydney_buildings <- buildings |> --> -->
<!-- <!--   filter(bbox$xmin > sydney_bbox[1], --> -->
<!-- <!--          bbox$ymin > sydney_bbox[2], --> -->
<!-- <!--          bbox$xmax < sydney_bbox[3], --> -->
<!-- <!--          bbox$ymax < sydney_bbox[4]) |> --> -->
<!-- <!--   select(id, geometry, height) |>  --> -->
<!-- <!--   collect() |> --> -->
<!-- <!--   st_as_sf(crs = 4326) |>  --> -->
<!-- <!--   mutate(height = ifelse(is.na(height), 4, height)) --> -->
<!-- <!-- ``` --> -->
<!-- <!-- ```{r rdeck sydney} --> -->
<!-- <!-- rdeck(map_style = mapbox_light(),  --> -->
<!-- <!--       initial_view_state = view_state( --> -->
<!-- <!--         center = c(151.186855, -33.871729), --> -->
<!-- <!--         zoom = 12, --> -->
<!-- <!--         bearing = -32.8, --> -->
<!-- <!--         pitch = 81 --> -->
<!-- <!--       )) |>  --> -->
<!-- <!--   add_polygon_layer( --> -->
<!-- <!--     data = sydney_buildings,  --> -->
<!-- <!--     name = "Sydney", --> -->
<!-- <!--     get_polygon = geometry,  --> -->
<!-- <!--     get_elevation = height,  --> -->
<!-- <!--     get_fill_color = scale_color_linear( --> -->
<!-- <!--       col = height, --> -->
<!-- <!--       palette = viridisLite::mako(100, direction = -1) --> -->
<!-- <!--     ), --> -->
<!-- <!--     extruded = TRUE,  --> -->
<!-- <!--     opacity = 0.5) --> -->
<!-- <!-- ``` --> -->

